Computing likelihoods under the SMC for a general class of ARGs.

### 2023-12-18 paper.tex
update time slice fig

[FIX ME: start is a mess. Need to find the right way in.]
Knowledge of the exact genealogy of a set of samples would make certain problems 
in population genetics (e.g.\ inferring recombination rates) trivial. 
Multitude of methods have started to 
demonstrate the value of attempts to capture the details of these intricately 
interwoven paths of inheritance referred to as an ancestral recombination graph (ARG).
Recent advancements: the inference of ARGs at population scale.
Yet even though potential of ARGs in population and statistical genetics has been 
demonstrated it remains unclear how to connect these reconstructed ARGs 
(as being approximate structures) to 
the stochastic/generative models that have lead us to study these datastructures 
in the first place.

@@

\begin{figure}[!ht]
\centering
\includegraphics[width=\textwidth]{figures/3-slices.png}
\caption{Simulated ARG under Hudson coalescent (left). $N_e$ varied was constant within each of the 3 time slices: $[(0:1000): 10 000, (1000:5000): 5000, (5000:\infty):10 000]$. Remaining parameters: L=10kb, 20 diploid samples, r=2.5e-8, 148 edges in ARG). A single $N_e$ value was estimated for each slice of the ARG as defined by the true population size change time points. The inferred posterior distribution (right) using the SMC-likelihood was obtained using a uniform prior. Number of edges per time slice: [68, 57, 23]. Locally unary nodes were omitted for clarity.}
\label{fig:3-arg-slices}
\end{figure}

>>>>

[FIX ME: introduction is very much work in progress.
start is a mess. Need to find the right way in.]
Knowledge of the exact genealogy of a set of samples would make certain problems 
in population genetics (e.g.\ inferring recombination rates) trivial.
Recent advancements in large-scale inference methods of the intricately 
interwoven paths of inheritance referred to as ancestral recombination graphs (ARGs)
\citep{rasmussen_genome-wide_2014, heine_bridging_2018, 
kelleher_inferring_2019, speidel_method_2019, rasmussen_espalier_2022, zhang_biobank-scale_2023}, 
are revolutionising the way we look at large genomic datasets.

Quite a number of studies have started to demonstrate the value of the recent ability to 
infer ARGs at scale by emphasizing the ability to rephrase many fundamental problems 
in genomics as questions on the transmission of genetic material from generation 
to generation [REFS].
Yet even though the potential of ARGs in population and statistical genetics has 
been demonstrated, the fact that these inferred ARGs lack explicit detail on 
recombination events means that it remains unclear how to connect these 
reconstructed ARGs to the stochastic/generative models that have lead us 
to study these datastructures in the first place.

@@

\begin{figure}[!ht]
\centering
\includegraphics[width=\textwidth]{figures/3-slices-whitespace.png}
\caption{Simulated ARG under Hudson coalescent (left). $N_e$ varied was constant within each of the 3 time slices: $[(0:1000): 10 000, (1000:5000): 5000, (5000:\infty):10 000]$. Remaining parameters: L=10kb, 20 diploid samples, r=2.5e-8, 148 edges in ARG). A single $N_e$ value was estimated for each slice of the ARG as defined by the true population size change time points. The inferred posterior distribution (right) using the SMC-likelihood was obtained using a uniform prior. Number of edges per time slice: [68, 57, 23]. Locally unary nodes were omitted for clarity.}
\label{fig:3-arg-slices}
\end{figure}
